They are Very Smart: Genetic-Algorithm-Based Pathfinding in Games

### 2018-12-12 paper.tex
Paper: Intro edits

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned
onslaught of enemies such as the fearsome Zombunny (see Figure 4). Zombunnies
are driven by a single goal: destruction of the player's home base. Zombunnies
spawn in groups of ten (separated by a slight delay) at the beginning of the
game. Subsequent waves spawn after all members from the previous wave have died.
Zombunnies use a genetic algorithm in conjunction with a recurrent neural
network to evolve a pathfinding algorithm between generations (waves), with the
long-term incentive of dealing damage to the player's base% (explain fitness
% criteria more concisely here)
. Zombunnies can be killed by either taking damage
from a player's defenses or through "starvation." A Zombunny will starve to
death if it does not take or deal damage for a given length of time. Through
this starvation mechanism, the Zombunnies which survive the longest are
ultimately those which interact the most with the player's defenses, and are
subsequently more likely to pass on attributes of their genome to the next
generation. Zombunnies have three characteristic attributes: health, damage, and
movement speed. All three of these stats are assigned upon spawning as dictated
by their genome. Pathfinding is controlled by a recurrent neural net, whose
weights are determined through a genetic algorithm between waves, which
optimizes for Zombunnies' ability to achieve their overarching goal of player
base destruction. Zombunnies can "see" nearby entities within a certain radius,
including other Zombunnies, player defenses, and the player's home base (when in
range). Based on the combination of visible inputs, Zombunnies will output a
chosen direction which they will then walk in. This algorithm updates on a
tick-by-tick basis. Zombunnies will damage structures they collide with. Because
evolution of the pathfinding neural net only happens between generations, near
the beginning of the game, Zombunnies will start off with relatively stupid
pathfinding which will typically cause the first few populations to starve to
death without dealing any significant damage. This provides the player with an
initial grace period during which they may build their initial defenses. As time
goes on, Zombunnies will evolve towards an optimal stat distribution of health,
damage, and speed, as well as in their ability to pathfind in such a way as to
thwart the player's defenses and end the game.

@@

\tavs~was initially inspired by Numantian Games' \textit{They Are Billions}, a game in which a player must build defenses in order to fend off a zombie apocalypse. The game features massive numbers of zombies. This project emerged from an interest in what \textit{They Are Billions} would be like if the zombies could actually think for themselves. The result is significantly pared down in comparison, as the focus has been on zombie evolution as opposed to player interactivity and enjoyability. However, similarities to \textit{They Are Billions} can be seen in the resource management system, the player's defense of a citadel, and need to fend off the faceless hordes. In contrast, \textit{They Are Billions} puts additional emphasis on world generation, player exploration, and narrative elements.

>>>>

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned
onslaught of enemies such as the fearsome Zombunny (see Figure 4). Zombunnies
are driven by a single goal: destruction of the player's home base. Zombunnies
spawn in groups of ten (separated by a slight delay) at the beginning of the
game. Subsequent waves spawn after all members from the previous wave have died.
Zombunnies use a genetic algorithm in conjunction with a recurrent neural
network to evolve a pathfinding algorithm between generations (waves), with the
ultimate goal of dealing damage to (and therefore destroying) the player's base
. Zombunnies can be killed by either taking damage from a player's defenses or
through "starvation." A Zombunny will starve to death if it does not take or
deal damage for a set moved distance. Through this starvation mechanism, the
Zombunnies which survive the longest are ultimately those which interact the
most with the player's defenses. Zombunnies have three
characteristic attributes: health, damage, and movement speed. All three of
these stats are assigned upon spawning as dictated by their genome. Pathfinding
is controlled by a recurrent neural net, whose topology and weight is encoded in
the genome alongside of the characteristic attributes. The neural net evolves on
a wave by wave basis, with those genomes that preform better being more likely
to survive and reproduce than their peers, optimizing for Zombunnies' ability to
achieve their overarching goal of player base destruction. Zombunnies can "see"
nearby entities within a certain radius, including other Zombunnies, player
defenses, and the player's home base (when in range). Based on the combination
of visible inputs, Zombunnies will output a chosen direction which they will
then walk in. This algorithm updates on a tick-by-tick basis. Zombunnies will
damage structures they collide with. Because evolution of the pathfinding neural
net only happens between generations, near the beginning of the game, Zombunnies
will start off with relatively stupid pathfinding which will typically cause the
first few populations to starve to death without dealing any significant damage.
This provides the player with an initial grace period during which they may
build their initial defenses. As time goes on, Zombunnies will evolve towards an
optimal stat distributions of health, damage, and speed, as well as in their
ability to pathfind in such a way as to thwart the player's defenses and end the
game.

@@

\tavs~was initially inspired by Numantian Games' \textit{They Are Billions}, a
game in which a player must build defenses in order to fend off a zombie
apocalypse. The game features massive numbers of zombies. This project emerged
from an interest in what \textit{They Are Billions} would be like if the zombies
could learn and evolve similarly to a real virus. The result is significantly pared down in
comparison, as the focus has been on zombie evolution as opposed to player
interactivity and enjoyment. However, similarities to \textit{They Are
  Billions} can be seen in the resource management system, the player's defense
of a citadel, and need to fend off the faceless hordes. In contrast,
\textit{They Are Billions} puts additional emphasis on world generation, player
exploration, and narrative elements.
